---
  title: "Boostrap ordination example"
  author: "Lan Huong Nguyen"
  date: "`r Sys.Date()`"
  output: rmarkdown::html_vignette
  vignette: >
  %\VignetteIndexEntry{analogues to plot ordination}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

##Loading / reshaping data
    
First we load the data from the phyloseq package.

```{r}
library("phyloseq")

data("GlobalPatterns")
GP = GlobalPatterns
wh0 = genefilter_sample(GP, filterfun_sample(function(x) x > 5), A = 0.5 * nsamples(GP))
GP1 = prune_taxa(wh0, GP)
# Pick low depth of the samples = 1000
GP1 = transform_sample_counts(GP1, function(x) 1e+03 * x/sum(x))
phylum.sum = tapply(taxa_sums(GP1), tax_table(GP1)[, "Phylum"], sum, na.rm = TRUE)
top5phyla = names(sort(phylum.sum, TRUE))[1:5]
GP1 = prune_taxa((tax_table(GP1)[, "Phylum"] %in% top5phyla), GP1)
human = get_variable(GP1, "SampleType") %in% c("Feces", "Mock", "Skin", "Tongue")
sample_data(GP1)$human <- factor(human)
```

## Boostrap Ordination

Define the distance we want to use, in this case a weighted UniFrac
```{r}
distWUnifrac <- function (D) {
  if (!all((dim(D) == dim(GP1@otu_table)))) GP1 <- t(GP1)
  rownames(D) <- taxa_names(GP1)
  colnames(D) <- sample_names(GP1)
  otu_table(GP1) <- otu_table(D, TRUE)
  phyloseq::UniFrac(GP1, weighted = TRUE, normalized = TRUE)
}
```


```{r}
bootOrd <- boot_ordination(GP1@otu_table, n = 20, method = "metaMDS", 
                          dist_method = distWUnifrac, common_depth = TRUE, 
                          replace_zero = FALSE, round = FALSE, scannf = F,
                          rows_annot = GP1@sam_data) 
mvarCenter <- bootOrd@center
mvarBoot <- mvar_boot2Table(bootOrd)

df <- cbind(mvarBoot@table[[1]]@coord, mvarBoot@table[[1]]@annotation)

plot_mvar(mvarBoot, col = "SampleType", shape = "human") + 
  stat_density_2d(data = df, aes(x = axis_1, y = axis_2, group = label, 
                                 col = SampleType, shape = human))
plot_mvar(mvarCenter, col = "SampleType", shape = "human")
```

## TO DO:

1. Implement the stat_density_2d layer for the plot_table
2. Allow options like size and pch in plot_mvar
3. How to add a center ordination plot with bigger size or different shape???
4. What is happening in "pco" from ade4 ?

```{r}
############################ Check what is happening ########################

origDist <- dist_method(GP1@otu_table)
orig_ord <- ade4::dudi.pco(origDist, scannf = F)

# if only distance object is supplied then what is orig_ord$li and what is orig_ord$co????
# shouldn't there be only one set of cooredinates???

# Warning message:
#   In ordi_method(X, ...) : Non euclidean distance
```
