---
title: "mvarVis d3 examples"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{plot_d3_examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This is an example from `MFA()` in the `FactoMineR` package, but done
using our wrapper. The first two columns are supplementary variables.

```{r}
library("mvarVis")
library("FactoMineR")
data(wine)
wine_mfa <- ordi(wine[, -c(1:2)], "MFA", wine[, 1:2], group = c(5,3,10,9,2),
                 type = rep("s",5), ncp = 3, graph = F)
```
We can plot the results d3. Hover over points to see their values in
the table that appears below the plot. Change the colors using the
selection box at the top left.

```{r, echo = F}
plot_mvar_d3(wine_mfa)
```
\
\
\
We can choose between `point`, `text`, and `arrow` for each layer
(though only one at a time, at least for now).

```{r}
plot_mvar_d3(wine_mfa, c("text", "arrow"))
```
\
\
\
If we want, we can have as many panels as we want. This splits the MFA
variables by group. The splitting code isn't pretty, but it shouldnt'
be too hard to write some helpers to streamline things.

```{r}
split_layer <- function(x, start, finish) {
  new("mvarLayer", coord = x@coord[start:finish, ],
      annotation = x@annotation[start:finish, ])
}

wine_quanti <- wine_mfa@table[["quanti.var"]]
wine_mfa_split <- new(
  "mvarTable",
  table = list(
    new("mvarLayer", coord = wine_mfa@table[["ind"]]@coord,
        annotation = wine_mfa@table[["ind"]]@annotation),
    split_layer(wine_quanti, 1, 5),
    split_layer(wine_quanti, 6, 8),
    split_layer(wine_quanti, 9, 18),
    split_layer(wine_quanti, 19, 27),
    split_layer(wine_quanti, 28, 29)
  ),
  eig = wine_mfa@eig
)

plot_mvar_d3(wine_mfa_split, c("point", rep("text", 5)))
```

We can also have just one panel.

```{r}
wine_pca <- ordi(wine[, -c(1:2)], "factominer_pca",
                 table_names = "ind",
                 ncp = 3, graph = F)
plot_mvar_d3(wine_pca)
```
\
\
\
Here's an MCA example. Notice we can customize width / height.

```{r}
data(hobbies)
hobbies_mca <- ordi(hobbies[, 1:18], method = "MCA",
                    rows_annot = hobbies[, 19:23], graph = F, ncp = 3)
```

```{r}
plot_mvar_d3(hobbies_mca, height = 700)
```
