---
title: "Analogues to ```plot_ordination```"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


The main inspiration for the ```mvarVis``` package comes from the
```plot_ordination``` command in package ```phyloseq```. Given the result of an
ordination, we should be able to plot various features onto the projected 
points easily. ```plot_ordination``` implements this idea in the case that
the ordination has features from a ```phyloseq``` object, but the idea 
can be applied more generally.

To work in this wider context, we currently have a 3-step pipeline

- Convert the result of an ```ade4``` or ```vegan``` ordination into a
class ```mvarTable```. Having this class means we don't have to rewrite 
plotting or annotation code for every kind of ordination package.
- Annotate the results of the ordination. By this we mean give covariates
that describe each of the projected points (like Phylum in the microbiome
context).
- Plot the annotated ```mvarTable``` object. The goal is to provide
flexibility in mapping aesthetic attributes without requiring too much
work from the user.

#```plot_ordination``` analogues

To get an idea for how this package works, here are some examples 
re-implementing the methods in ```plot_ordination``` from the
[phyloseq documentation](http://joey711.github.io/phyloseq/plot_ordination-examples).

##Loading / reshaping data

First we load the data from the linked page.


```{r}
library("ape")
library("ade4")
library("phyloseq")
library("vegan")
library("mvarVis")
```

```{r}
data("GlobalPatterns")
GP = GlobalPatterns
wh0 = genefilter_sample(GP, filterfun_sample(function(x) x > 5), A = 0.5 * nsamples(GP))
GP1 = prune_taxa(wh0, GP)
GP1 = transform_sample_counts(GP1, function(x) 1e+06 * x/sum(x))
phylum.sum = tapply(taxa_sums(GP1), tax_table(GP1)[, "Phylum"], sum, na.rm = TRUE)
top5phyla = names(sort(phylum.sum, TRUE))[1:5]
GP1 = prune_taxa((tax_table(GP1)[, "Phylum"] %in% top5phyla), GP1)
human = get_variable(GP1, "SampleType") %in% c("Feces", "Mock", "Skin", "Tongue")
sample_data(GP1)$human <- factor(human)
```

### ```vegan``` methods

```{r}
gp1_nmds <- ordi(GP1@otu_table, GP1@tax_table, GP1@sam_data, method = "metaMDS")
plot_mvar(gp1_nmds, col = "Phylum", shape = "human")

# DCA
gp1_dca <- ordi(GP1@otu_table, GP1@tax_table, GP1@sam_data, method = "decorana")
plot_mvar(gp1_dca, col = "Phylum", shape = "human")
```

### ```ade4``` methods

There is an analogous function for ```ade4```. Note that we can pass in ```ade4```
arguments through ```ordi``` (as in the ```scannf = F``` below).

```{r}
# DPCoA
patristicDist <- cailliez(as.dist(cophenetic.phylo(phy_tree(GP1))))
gp1_dpcoa <- ordi(list(GP1@otu_table, patristicDist), GP1@tax_table, GP1@sam_data, 
                  method = "dpcoa", scannf = F)
plot_mvar(gp1_dpcoa, "point-text", col = "Phylum")

# PCA
gp1_pca <- ordi(GP1@otu_table, GP1@tax_table, GP1@sam_data, method = "pca", 
                scannf = F, nf = 10)
plot_mvar(gp1_pca, col = "Phylum", facet_vector = c("SampleType", "Phylum"))

# MDS
gp1_mds <- ordi(GP1@otu_table, GP1@tax_table, GP1@sam_data, method = "pco", 
                dist_method = "euclidean", scannf = F)
plot_mvar(gp1_mds, col = "Phylum", shape = "human")
```

I haven't implemented the pcoa example yet, since this needs a conversion
from an ```ape``` specific format.

