---
title: "Analogues to ```plot_ordination```"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


The main inspiration for the ```mvarVis``` package comes from the
```plot_ordination``` command in package ```phyloseq```. Given the result of an
ordination, we should be able to plot various features onto the projected 
points easily. ```plot_ordination``` implements this idea in the case that
the ordination has features from a ```phyloseq``` object, but the idea 
can be applied more generally.

To work in this wider context, we currently have a 3-step pipeline

- Convert the result of an ```ade4``` or ```vegan``` ordination into a
class ```mvarTable```. Having this class means we don't have to rewrite 
plotting or annotation code for every kind of ordination package.
- Annotate the results of the ordination. By this we mean give covariates
that describe each of the projected points (like Phylum in the microbiome
context).
- Plot the annotated ```mvarTable``` object. The goal is to provide
flexibility in mapping aesthetic attributes without requiring too much
work from the user.

#```plot_ordination``` analogues

To get an idea for how this package works, here are some examples 
re-implementing the methods in ```plot_ordination``` from the
[phyloseq documentation](http://joey711.github.io/phyloseq/plot_ordination-examples).

##Loading / reshaping data

First we load the data from the linked page.


```{r}
library("ape")
library("ade4")
library("phyloseq")
library("vegan")
library("mvarVis")
```

```{r}
data("GlobalPatterns")
GP = GlobalPatterns
wh0 = genefilter_sample(GP, filterfun_sample(function(x) x > 5), A = 0.5 * nsamples(GP))
GP1 = prune_taxa(wh0, GP)
GP1 = transform_sample_counts(GP1, function(x) 1e+06 * x/sum(x))
phylum.sum = tapply(taxa_sums(GP1), tax_table(GP1)[, "Phylum"], sum, na.rm = TRUE)
top5phyla = names(sort(phylum.sum, TRUE))[1:5]
GP1 = prune_taxa((tax_table(GP1)[, "Phylum"] %in% top5phyla), GP1)
human = get_variable(GP1, "SampleType") %in% c("Feces", "Mock", "Skin", "Tongue")
sample_data(GP1)$human <- factor(human)
```

### ```vegan``` methods

This is a ```vegan_to_mvar``` object that should make it easy
to use this package with ```vegan```.

```{r}
gp1_nmds <- vegan::metaMDS(GP1@otu_table, dist = "bray")
gp1_nmds_mvar <- vegan_to_mvar(gp1_nmds)
gp1_nmds_mvar <- mvar_annotate(gp1_nmds_mvar, list(GP1@tax_table, GP1@sam_data))

sample_plot_opts <- list(aes_list = list(color = "Phylum"))
variable_plot_opts <- list(aes_list = list(label = "SampleType"),
                           layers_list = list(point = FALSE, text = TRUE))
plot_mvar(gp1_nmds_mvar, list(sample_plot_opts, variable_plot_opts))

# DCA
gp1_dca <- vegan::decorana(GP1@otu_table)
gp1_dca_mvar <- vegan_to_mvar(gp1_dca)
gp1_dca_mvar <- mvar_annotate(gp1_dca_mvar, list(GP1@tax_table, GP1@sam_data))
plot_mvar(gp1_dca_mvar, list(sample_plot_opts, variable_plot_opts))


# cca
gp1_cca <- vegan::cca(GP1@otu_table)
gp1_cca_mvar <- vegan_to_mvar(gp1_cca)
gp1_cca_mvar <- mvar_annotate(gp1_cca_mvar, list(GP1@tax_table, GP1@sam_data))
p <- plot_mvar(gp1_cca_mvar, color_vars = list("Phylum", NULL),
               arrow_ix = numeric(0), text_ix = 2, text_vars = list(NULL, "SampleType"),
               shape_vars = list(NULL, "human"))
p
# rda
gp1_rda <- vegan::rda(GP1@otu_table)
gp1_rda_mvar <- vegan_to_mvar(gp1_rda)
gp1_rda_mvar <- mvar_annotate(gp1_rda_mvar, list(GP1@tax_table, GP1@sam_data))
p <- plot_mvar(gp1_rda_mvar, color_vars = list("Phylum", NULL),
               arrow_ix = numeric(0), text_ix = 2, text_vars = list(NULL, "SampleType"),
               shape_vars = list(NULL, "human"))
p
```

### ```ade4``` methods

There is an analogous function for ```ade4```.

```{r}
# DPCoA
patristicDist <- cailliez(as.dist(cophenetic.phylo(phy_tree(GP1))))
gp1_dpcoa <- dpcoa(data.frame(GP1@otu_table), patristicDist, scannf = F, nf = 2)
gp1_dpcoa_mvar <- ade4_to_mvar(gp1_dpcoa, tables_to_include = c("l1", "l2"))
gp1_dpcoa_mvar <- mvar_annotate(gp1_dpcoa_mvar, list(GP1@tax_table, GP1@sam_data))
plot_mvar(gp1_dpcoa_mvar, color_vars = list("Phylum", NULL), arrow_ix = numeric(0),
          text_ix = 2, text_vars = list(NULL, "SampleType"), shape_vars = list(NULL, "human"))

# MDS
gp1_mds <- dudi.pco(dist(GP1@otu_table), scannf = F, nf = 2)
gp1_mds_mvar <- ade4_to_mvar(gp1_mds)
gp1_mds_mvar <- mvar_annotate(gp1_mds_mvar, list(GP1@tax_table, GP1@sam_data))
plot_mvar(gp1_mds_mvar, color_vars = list("Phylum", NULL), arrow_ix = numeric(0),
          text_ix = 2, text_vars = list(NULL, "SampleType"), shape_vars = list(NULL, "human"))
```

I haven't implemented the pcoa example yet, since this needs a conversion
from an ```ape``` specific format.
